On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-10-13 Manuscript/manuscript.tex
went over discussion

\section{Concluding Remarks}\label{sec:concludingremarks}

We did not consider data from the women's 100 meter hurdle in the analysis.
In addition to the concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 was at best only
slightly faster relative to other championships going back to 2003.
Likewise, the rank-based comparisons resulted in p-values
that were not drastically significant after adjusting for multiple tests.
The lack of evidence from the women's data makes it 
harder to conclude that the sensors used at the 2022 championships were faulty, 
although it is unknown if the sensors for the men's 110 meter hurdle and women's
100 meter hurdle were the same. On the other hand, this may be related to 
another debated question regarding whether the reaction time barrier should be 
the same for men and women.

It is confusing as to why the reaction time barrier for men and women 
is the same despite numerous articles \citep[e.g.,][]{lipps2011implications, 
babicc2009reaction, panoutsakopoulos2020gender} stating differently. 
According to these authors, World Athletics is ignoring differences in gender by 
setting the reaction time barrier to be equal. 
If there are gender differences and 0.1 seconds is considered a fair time for 
women, then 0.1 seconds could not be considered fair for men, as the probability 
of observing reaction times lower than 0.1 would be elevated.
\citet{brosnan2017effects} advocated for gender based reaction time barriers 
which seems logical if there are biological
differences in reaction times between men and women.

The first part of this paper showed that athletes' reaction times at the 2022 
World Track and Field Championships were generally faster than their reaction 
times at other competitions. Table~\ref{tab:Clusrankresults} shows multiple 
significant p-values that indicate differences in average reaction times for 
athletes who competed at the 2022 World Track and Field Championship and another
competition. Additionally, this paper has shown that a reaction time of 0.1 
seconds is not as extreme as one may think, and that based on the probabilities
in Table~\ref{tab:Sim_time}, World Athletics may consider changing the 
disqualification barrier to 0.8 seconds. If they were to do so, they would 
allow athletes like Allen to be able to react faster without fear of being 
disqualified. Taken together, the results show that 2022 was an outlier and 
despite being disqualified, Allen's time was not so extreme.

>>>>

\section{Discussion}\label{sec:concludingremarks}

We did not consider data from the women's 100 meter hurdle in the analysis.
In addition to the concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 women were at best only
slightly faster relative to other championships going back to 2003.
Likewise, the rank-based comparisons resulted in p-values
that were not drastically significant after adjusting for multiple tests.
The lack of evidence from the women's data makes it 
harder to conclude that the sensors used at the 2022 championships were faulty, 
although it is unknown if the sensors for the men's 110 meter hurdle and women's
100 meter hurdle were the same. On the other hand, this may be related to 
another debated question regarding whether the reaction time barrier should be 
the same for men and women.

The uniformity of the reaction time barrier for both men and women is
perplexing, especially in light of numerous studies that suggest a divergence in
their respective reaction times \citep[e.g.,][]{lipps2011implications,
  babicc2009reaction, panoutsakopoulos2020gender}. These authors posit that
World Athletics may be overlooking inherent gender differences by establishing
an equal reaction time barrier. If gender disparities do exist and 0.1 seconds
is deemed a fair threshold for women, it logically follows that the same cannot
be deemed fair for men, given that the probability of observing sub-0.1-second
reaction times would be inherently higher. \citet{brosnan2017effects} propounds
the adoption of gender-specific reaction time barriers, a stance that appears
logical when considering biological distinctions in reaction times between
genders.

This paper aspires to offer a statistical lens through which to examine Allen's
disqualification at the 2022 World Track and Field Championships, rather than
delivering a conclusive judgment regarding potential equipment malfunction. Our
findings reveal that athletes' reaction times at the event were, in general,
faster than those recorded at other competitions, with
Table~\ref{tab:Clusrankresults} presenting multiple significant p-values that
highlight disparities in average reaction times between the 2022 Championship
and other competitions. Further, our GLMM results indicate that a reaction time
of 0.1 seconds may not be as exceptional as commonly believed. According to the
probabilities in Table~\ref{tab:Sim_time}, World Athletics might contemplate
adjusting the disqualification barrier to 0.8 seconds, thereby enabling athletes
like Allen to react more swiftly without the apprehension of
disqualification. In summary, while the results designate 2022 as an anomalous
year, Allen's time, despite resulting in disqualification, may not be
categorically extreme.
